The Gravettian-Solutrean transition in westernmost Iberia: new data from the sites of Vale Boi and Lapa do Picareiro

### 2020-04-22 analysis/paper/paper.Rmd
methods changed

The materials used for this analysis include all lithic artifacts from Vale Boi's layers 5 and 4E of the Terrace area and levels U and T from Lapa do Picareiro.

The present study adopted an attribute analysis methodology, which aims to recover morphological and metrical attributes of technological classes. The attributes analyzed followed those previously used in European and Portuguese contexts for the Upper Paleolithic [see @bicho1992; @zilhao1997; @almeida2000; @cascalheira2013].

For this study, all artifacts and their technological and morphological attributes were recorded using two different databases, thus dividing the analysis in two phases: basic and complete attribute-based analysis.

In both cases, data collection was done through E4, a software developed by Dibble and McPherron (2003), available for download at http://www.oldstoneage.com/software/default.shtml. The software allows the user to program a database file which filters variables through conditions, defined by previous choices and attributes, which allows for better control of the database and homogeneous results. After recording, all information is gathered in one single database in an Access type output.

@@

The several attributes analysed in this database follow those present in specialized literature, such as @brezillon1968 and @tixier1980, paired with the methodologies used in Upper Paleolithic lithic attribute analysis works, such as @bicho1992, @zilhao1997 or @almeida2000.

The complete Data Dictionary with all variables, attributes, values and description, with reference to the consulted literature can be found in the supplementary material.

The measurements were registered equally for debitage products and cores, including maximum length, width (proximal, mesial and distal), thickness, platform width and thickness and platform angle according to @dibble1997.

After data collection, the databases were imported into R environment, where the information was processed through the creation of descriptive statistical analysis and writing of this article. We include the entire R code used for all the analysis and visualizations contained in this paper is in our online research compendium at https://doi.org/. To produce those files, we followed the procedures described by Marwick et al. (XXXX) for the creation of research compendiums to enhance the reproducibility of research. To enable maximum re-use, our code is released under the MIT license, our data as CC-0, and our figures as CC-BY (for more information, see Marwick XXXX).

@@

## Vale Boi

@@

### Technology

@@

## Lapa do Picareiro

>>>>

The materials used for this analysis include all lithic artifacts from Vale Boi's layers 5 and 4E of the Terrace area and levels U and T from Lapa do Picareiro (see below).

For lithic analysis we adopted an attribute-based methodology, which aims to describe morphological and metrical attributes of technological classes. The attributes analyzed followed those present in specialized literature, such as @brezillon1968 and @tixier1980, paired with the methodologies used in Upper Paleolithic lithic attribute analysis works, such as @bicho1992, @zilhao1997 or @almeida2000.

Artifacts and their techno-morphological attributes were recorded using two different databases, thus dividing the analysis in two phases: basic and complete attribute-based analysis.

@@

Data collection was done through E4, a software developed by Dibble and McPherron (2003). The complete Data Dictionary with all variables, attributes, values and description, with reference to the consulted literature is available in our research compendium at https://doi.org/xxxx.

After data collection, the databases were imported into R environment, where the information was processed through the creation of descriptive statistical analysis and writing of this article. We include the entire R code used for all the analysis and visualizations contained in this paper is in our online research compendium. To produce those files, we followed the procedures described by Marwick et al. (XXXX) for the creation of research compendiums to enhance the reproducibility of research. To enable maximum re-use, our code is released under the MIT license, our data as CC-0, and our figures as CC-BY (for more information, see Marwick XXXX).

@@

# Vale Boi

@@

## Lithic technology

@@

# Lapa do Picareiro
